The irreducibility of primes can be used by the following lemma.
\begin{lemma}[Euclid's Lemma in $\Qx$]
Let $p(x),f(x),g(x)\in\Qx$, and suppose $p(x)$ is prime. If $p(x)\mid f(x)g(x)$, then $p(x)\mid f(x)$ or $p(x)\mid g(x)$.
\end{lemma}

\begin{proof}
Suppose $p(x)\mid f(x)g(x)$. If $p(x)\mid f(x)$, then we are done. So assume $p(x)\nmid f(x)$.

Let $d(x)=\gcd(p(x),f(x))$. Since $d(x)\mid p(x)$ and $p(x)$ is prime, $d(x)$ is either a unit or an associate of $p(x)$.

If $d(x)$ were an associate of $p(x)$, then $p(x)\mid d(x)$. Since $d(x)\mid f(x)$, this would imply $p(x)\mid f(x)$, contradicting our assumption. Therefore $d(x)$ is a unit.

Since gcds are chosen to be monic, the only monic unit is $1$. Hence $d(x)=1$. By Bezout's Identity, there exist $a(x),b(x)\in\Qx$ such that
\[
a(x)p(x)+b(x)f(x)=1.
\]
Multiplying both sides by $g(x)$ gives
\[
a(x)p(x)g(x)+b(x)f(x)g(x)=g(x).
\]
Now $p(x)$ divides $a(x)p(x)g(x)$ because that term has a factor of $p(x)$. Also, $p(x)$ divides $b(x)f(x)g(x)$ because $p(x)\mid f(x)g(x)$. Therefore $p(x)$ divides their sum, so $p(x)\mid g(x)$.

Thus $p(x)\mid f(x)$ or $p(x)\mid g(x)$.
\end{proof}
This helpful corollary extends this to the greatest common divisor and allows us to multiply them. If a number is relatively prime to two numbers, it must be relatively prime to their product.
\begin{corollary}
    If $f(x), g(x), h(x) \in \qx$, and $\gcd(f, g) = 1$ as well as $\gcd(f, h) = 1$ then $\gcd(f, gh) = 1$.
\end{corollary}
\begin{proof}
    Assume for the sake of contradiction that there exists $a(x) \in \qx$, $\|a(x)\| > 1$ such that $\gcd(f, gh) = a$. Thus, $a \mid f$ and $a \mid gh$. We know that $a(x)$ must have a prime factor. Define it as $p(x)$. Now, we know that \[
    p(x) \mid f(x) \qquad p(x) \mid g(x)h(x)
    \]
    By the Fundamental Lemma, $p(x) \mid g(x)$ or $p(x) \mid h(x)$. Without loss of generality, let $p(x) \mid g(x)$. Thus, since $p(x) \mid f(x)$, $\gcd(f(x), g(x)) \geq \|p(x)\| > 1$. However, this is a contradiction. Thus, $\gcd(f, gh) = 1$
\end{proof}

Now, we can extend these lemmas to a more general case that proves if a number divides a product, not all numbers in the product can be relatively prime to that number.
\begin{theorem}
    Let $f(x),g_i(x)\in\Qx$, and let $f(x) \mid \prod_{i=1}^n g_i(x)$. Then, there exists $k 
    \leq n$ such that $\gcd(f(x), g_k(x)) \neq 1$. 
\end{theorem}
\begin{proof}
    Prove that if $\gcd(f(x), g_i(x)) = 1$ for all $i \leq n$, then $\gcd(f(x), \prod_{i=1}^n g_i(x)) = 1$.
    Let $S= {j \in \pZ | \gcd(f(x), \prod_{i=1}^n g_i(x)) \neq 1}.$. For the sake of contradiction, S is not empty. Then, since $S\subseteq \pZ$, by the well-ordering principle, there exists a least element $l \in S$. Note that if $j=1$, then \[
        \gcd(f(x), \prod_{i=1}^1 g_i(x))  = \gcd(f(x), g_1(x))  =  1
    \]which is a contradiction. Thus, since $1 \notin S$, $l>1$ and $l-1 \notin S$. So, we know that \[
    \gcd(f(x), \prod_{i=1}^{l-1} g_i(x)) =  1
    \] Since $\gcd(f(x), \prod_{i=1}^{l-1} g_i(x)) =  1$ and $\gcd(f(x), g_l(x)) =  1$, $\gcd(f(x), g_l(x)\prod_{i=1}^{l-1} g_i(x)) =\gcd(f(x), \prod_{i=1}^{l} g_i(x)) =  1$. However, since $l \in S$, this is a contradiction. Thus, $S$ must be empty.
    As $\gcd(f(x), \prod_{i=1}^n g_i(x)) = 1$ is a contradiction to the fact that $f(x) \mid \prod_{i=1}^n g_i(x)$,  $\gcd(f(x), g_i(x)) $ cannot equal 1 for all $i \leq n$.
\end{proof}



