Now we can formally define our rational polynominals.
\subsection{Construction}
\begin{definition}[The polynomial ring $\Qx$]
We define a polynomial $f(x) \in \Qx$ with degree $n$ if it can be written as
\[
f(x)=\sum_{0}^{n} a_i x^i
\]
where $n\in\mathbb{Z}_{\geq0}$ and each coefficient $a_i$ lies in $\mathbb{Q}$ and $a_n \neq 0$. For convenience, $f(x)$ can also be notated as $f$.
\\ Define two operations $+, \m$ such that, given $f(x) = \sum_{i=0}^{n} a_i x^i$ and $g(x) = \sum_{i=0}^{m} b_i x^i $
\begin{itemize}
    \item $+$: $\begin{cases}
        m>n &  f(x) + g(x) = \sum_{i=0}^{n} (a_i +b_i)x^i + \sum_{i=n+1}^{m} (b_i)x^i \\ 
        n<m & f(x) + g(x) = \sum_{i=0}^{m} (a_i +b_i)x^i + \sum_{i=m+1}^{n} (a_i)x^i
    \end{cases}$\\
    Since $\mathbb{Q}$ is closed under addition, each
$a_i+b_i$ is rational. Hence $f(x)+g(x)\in\Qx$.
    \item $\times$: By Multiplied Sums, $f(x)\cdot g(x) = \sum_{i=0}^{n} \sum_{j=0}^{m} a_i b_j x^i x^j$
\end{itemize}
\end{definition}

\subsection{Ring Axioms}
The rational polynominals also follow the ring axioms.
\noindent\textbf{Commutativity of Addition: } 
Addition is commutative because each coefficient is added in
$\mathbb{Q}$, and addition in $\mathbb{Q}$ is commutative.
\noindent\textbf{Associativity of Addition: } 
Addition is associative because each coefficient is added in
$\mathbb{Q}$, and addition in $\mathbb{Q}$ is associative.
\textbf{Commutativity of Multiplication: }
Multiplication is commutative because the coefficient of $x^k$ in $f(x)g(x)$
comes from all products $a_i b_j$ with $i+j=k$, while the coefficient of $x^k$
in $g(x)f(x)$ comes from all products $b_j a_i$ with $j+i=k$. These are the
same products, since multiplication in $\mathbb{Q}$ is commutative.

\noindent\textbf{Associativity of Multiplication: }
Multiplication is associative because the coefficient of $x^k$ in
$(f(x)g(x))h(x)$ comes from all products
\[
(a_i b_j)c_r
\]
with $i+j+r=k$, while the coefficient of $x^k$ in $f(x)(g(x)h(x))$ comes from
all products
\[
a_i(b_j c_r)
\]
with $i+j+r=k$. These are equal because multiplication in $\mathbb{Q}$ is
associative.
 
\noindent\textbf{Distributivity: }
Distributivity holds coefficient by coefficient. Put missing coefficients of
$g(x)$ and $h(x)$ equal to $0$. Then the coefficient of $x^k$ in
$f(x)(g(x)+h(x))$ uses terms of the form
\[
a_i(b_j+c_j).
\]
Since
\[
a_i(b_j+c_j)=a_i b_j+a_i c_j
\]
in $\mathbb{Q}$, we get
\[
f(x)(g(x)+h(x))=f(x)g(x)+f(x)h(x).
\]

Finally, $0_{\Qx}$ has constant coefficient $0$, while $1_{\Qx}$ has constant
coefficient $1$. Since $0\neq1$ in $\mathbb{Q}$,
\[
0_{\Qx}\neq1_{\Qx}.
\]
\noindent\textbf{Additive Identity: }
The zero polynomial, $0_{\Qx} = 0$, is the additive identity, as 
\[
f(x)+0_{\Qx}
=(a_0+0)+\sum_{i=1}^n a_ix^i
\]
 
\noindent\textbf{Negatives: }
Define $-f(x)$ as:
\[
-f(x)=\sum_{i=0}^n (-a_i)x^i
\]
This is the additive inverse of $f(x)$ because $a_i+(-a_i)=0$ for every $i$.

\noindent\textbf{Multiplicative Identity: }
Define $1_{\Qx} = 1$ as the multiplicative identity. In the product
$1_{\Qx}f(x)$, the coefficient of $x^i$ is $1\cdot a_i=a_i$. Hence
\[
1_{\Qx}f(x)=f(x).
\]


% For each $k$ with $0\leq k\leq m+n$, let //idk how useful knowing this is
% \[
% P_k=\{(i,j)\in\mathbb{Z}^2:0\leq i\leq m,\ 0\leq j\leq n,\ i+j=k\}.
% \]
% Write
% \[
% P_k=\{(i_1,j_1),\ldots,(i_t,j_t)\}.
% \]
% Define
% \[
% d_k=a_{i_1}b_{j_1}+\cdots+a_{i_t}b_{j_t}.
% \]
% Each $d_k$ is rational, since it is made from finitely many additions and
% multiplications in $\mathbb{Q}$. Therefore
% \[
% f(x)g(x)=d_0+d_1x+\cdots+d_{m+n}x^{m+n}
% \]
% is in $\Qx$.
\subsection{Units}
Just like integer rings, the polynominals have units as well that follow the same definition.
\begin{definition}
    Define $u(x) \in \qx$ as a \textbf{unit} if there exists $v(x) \in \qx$ s.t. $u(x)v(x) = 1$. The set of units in $\qx$ is denoted $\qx^{\times}$.
\end{definition}
\begin{lemma}
    $u(x) \in \qx^{\times} \; \Leftrightarrow \; u(x) \in \bQ$
\end{lemma}
\begin{proof}
    First, prove $u(x) \in \bQ \; \Rightarrow  \; u(x) \in \qx^{\times}$. This is a collorary of the Multiplicative Inverse property of $\bQ$. 
    Second, prove $u(x) \in \qx^{\times} \; \Rightarrow \; u(x) \in \bQ$. Define $u(x)=\sum_{i=0}^{n} a_i x^i$ and $v(x)=\sum_{j=0}^{m} b_j x^j$, with $n$ and $m$ the degrees of $u$ and $v$ respectively, then 
        \begin{align*}
        u(x)v(x)&=\sum_{i=0}^{n} \sum_{j=0}^{m} a_i b_j x^i x^j=\sum_{i=0}^{n} \sum_{j=0}^{m} a_i b_j x^{i+j}
        \end{align*}
        by Multiplied Sums. Notice that $i+j = m+n$ only once. However, $a_mb_n$ are both nonzero, so $a_mb_n \neq 0$. Thus, we have a nonzero term $a_mb_nx^{m+n} \neq 1$.  
\end{proof}
Notice that because units in $\qx$ are constants, this makes every polynominal have an equivalent one that is monic.

\subsection{The Norm}
We need a method to define size of a polynominal. The natural choice is its degree; however, this is not multiplicative, and norms are more helpful when they are multiplicative.
\begin{definition}
    For $f(x) \in \Qx / \{0\}$, let the \textbf{norm}, or $\| f(x)\|$ be equal to \[
    \| f(x) \|  = 2^{\deg{f(x)}}
    \]
    Define $\|0\|$ to be 0.
with the following properties:
\begin{itemize}
    \item $\| f(x)\| \cdot\| g(x)\| = \| f(x)g(x)\|$ 
    \begin{proof}
        By Multiplied Sums, if $f(x)=\sum_{i=0}^{n} a_i x^i$ and $g(x)=\sum_{j=0}^{m} b_j x^j$, with $n$ and $m$ the degrees of $f$ and $g$ respectively, then 
        \begin{align*}
        f(x)g(x)&=\sum_{i=0}^{n} \sum_{j=0}^{m} a_i b_j x^i x^j=\sum_{i=0}^{n} \sum_{j=0}^{m} a_i b_j x^{i+j}
        \end{align*}
        Notice that to find the degree of $f(x)g(x)$, we take the maximum value of $i+j$ which is their respective maximums, $n+m$, so:
        \[
        \|f(x)g(x)\|=2^{n+m}=2^n2^m = \|f(x)g(x)\|
        \]
    \end{proof}
    \item For any $c(x) \in \Qx ^ {\times}$, $\|c(x)\| = 1$.
    \begin{proof} Since $\deg{c(x)} = 0$, $\|c(x)\|=2^0 = 1$.\end{proof}
    \item $\| f(x)\| \in \bZ \cup \{0\}$. Since $2 \in \bZ^+$, $2^n$ for $n\in \bZ^+$ is always positive by Appendix B, the norm $\in \bZ \cup \{0\}$
\end{itemize}
\end{definition}

We want units (and only units) to preserve norm.
\begin{lemma}
    $\|f(x)g(x)\| = \|f(x)\| \Leftrightarrow g(x) \in \qx^{\times}$ 
\end{lemma}
\begin{proof}
    \begin{itemize}
        \item  $ g(x) \in \qx^{\times} \Rightarrow \|f(x)g(x)\| = \|f(x)\|$
        \begin{proof}
            Since $\|g(x)\| = 1$, $\|f(x)g(x)\| = \|f(x)\| \m 1 = \|f(x)\|$.
        \end{proof}
        \item $\|f(x)g(x)\| = \|f(x)\| \Rightarrow g(x) \in \qx^{\times}$ 
        \begin{proof}
            Since $\|f(x)g(x)\| = \|f(x)\| \|g(x)\|$. As $ \|f(x)\|, \|g(x)\| \in \bZ^+$, if $\|f(x)g(x)\| = \|f(x)\|$, $\|g(x)\| = 1$. Thus, $g(x) \in \qx^{\times}$.
        \end{proof}
    \end{itemize}
\end{proof}

In addition, $\qx$ is an integral domain.
\begin{theorem}[$\Qx$ has no zero divisors]
If $f(x),g(x)\in\Qx$ and
\[
f(x)g(x)=0,
\]
then $f(x)=0$ or $g(x)=0$.
\end{theorem}
\begin{proof}
Suppose, for contradiction, that
\[
f(x)g(x)=0
\]
with $f(x)\neq0$ and $g(x)\neq0$.

Since $g(x)\neq0$, by the norm lemma, %%todo: where is this lemma
\[
\|f(x)\|\leq \|f(x)g(x)\|.
\]
Since $f(x)\neq0$, $\|f(x)\|$ is positive.
So $\|f(x)g(x)\|$ is positive. In particular, $f(x)g(x)\neq0$, since the norm
is only positive for nonzero polynomials.

This contradicts $f(x)g(x)=0$. Therefore at least one of $f(x)$ and $g(x)$
must be zero.
\end{proof}

\begin{corollary}[Cancellation]
If $h(x)\neq0$ and
\[
h(x)f(x)=h(x)g(x),
\]
then $f(x)=g(x)$.
\end{corollary}
\begin{proof}
Subtract the right side from the left:
\[
h(f-g)=0.
\]
Since $h\neq0$ and $\Qx$ has no zero divisors, we must have
$f-g=0$.  Thus $f=g$.
\end{proof}

\subsection{Basic Arithmetic}
Almost all the integer properties have analogous properties in $\qx$.
\begin{lemma}[Uniqueness of zero and one]
The additive identity and multiplicative identity in $\Qx$ are unique.
In particular, if $f(x)+z(x)=f(x)$ for every $f(x)\in\Qx$, then
$z(x)=0$.  Similarly, if $f(x)u(x)=f(x)$ for every $f(x)\in\Qx$, then
$u(x)=1$.
\end{lemma}

\begin{proof}
Suppose $z$ acts like an additive identity for $f$, so $f+z=f$.  Add $-f$ to
both sides:
\[
(-f)+(f+z)=(-f)+f.
\]
By associativity and the definition of additive inverse, this becomes
\[
0+z=0.
\]
Since $0$ is the additive identity, $z=0$.

Now suppose $u$ acts like a multiplicative identity for every polynomial.  Test
this on the constant polynomial $1$:
\[
1\cdot u=1.
\]
But $1\cdot u=u$.  Hence $u=1$.
\end{proof}

\begin{lemma}[Uniqueness of additive inverses]
If $u(x)$ and $v(x)$ are both additive inverses of $f(x)$, then $u(x)=v(x)$.
\end{lemma}

\begin{proof}
Assume
\[
f+u=0
\qquad\text{and}\qquad
f+v=0.
\]
Then $f+u=f+v$.  Add $-f$ to both sides:
\[
(-f)+(f+u)=(-f)+(f+v).
\]
By associativity,
\[
((-f)+f)+u=((-f)+f)+v.
\]
Thus
\[
0+u=0+v,
\]
so $u=v$.
\end{proof}

\begin{lemma}[Zero and negatives]
For all $f(x),g(x)\in\Qx$,
\[
f(x)\cdot0=0,\qquad -(-f(x))=f(x),\qquad (-f(x))(-g(x))=f(x)g(x).
\]
\end{lemma}

\begin{proof}
First,
\[
f\cdot0=f(0+0)=f\cdot0+f\cdot0.
\]
Adding the additive inverse of $f\cdot0$ to both sides gives $f\cdot0=0$.

Next, since
\[
(-f)+f=0,
\]
the polynomial $f$ is an additive inverse of $-f$.  By uniqueness of additive
inverses,
\[
-(-f)=f.
\]

For the product of negatives, first observe that
\[
fg+(-f)g=(f+(-f))g=0\cdot g=0.
\]
Therefore $(-f)g=-(fg)$.  Applying this with $g$ replaced by $-g$ gives
\[
(-f)(-g)=-(f(-g)).
\]
Also,
\[
fg+f(-g)=f(g+(-g))=f\cdot0=0,
\]
so $f(-g)=-(fg)$.  Hence
\[
(-f)(-g)=-(-(fg))=fg.
\]
\end{proof}

\begin{lemma}
Let $f(x),g(x)\in\Qx$ with $f(x)\neq0$ and $g(x)\neq0$. Then
\[
\|f(x)\|\leq \|f(x)g(x)\|.
\]
\end{lemma}

\begin{proof}
Let
\[
\deg(f)=m
\qquad\text{and}\qquad
\deg(g)=n.
\]
Since $g(x)\neq0$, we have $n\geq0$. Also, since $f(x)\neq0$ and
$g(x)\neq0$, the leading coefficient of $f(x)g(x)$ is the product of the
leading coefficients of $f(x)$ and $g(x)$, which is nonzero in $\mathbb{Q}$.
Thus
\[
\deg(fg)=\deg(f)+\deg(g)=m+n.
\]
Therefore
\[
\|f(x)g(x)\|=2^{\deg(fg)}=2^{m+n}.
\]
Since $n\geq0$,
\[
2^m\leq2^{m+n}.
\]
So
\[
\|f(x)\|=2^m\leq2^{m+n}=\|f(x)g(x)\|.
\]
\end{proof}


\subsection{Divisibility}
Only with divisibility can we define fundamental lemmas and primes.
\begin{definition}[Divisibility]
For $f(x),g(x)\in\Qx$, we say $f(x)$ divides $g(x)$, written
\[
f(x)\mid g(x),
\]
if there exists $q(x)\in\Qx$ such that
\[
g(x)=f(x)q(x).
\]
\end{definition}
The greatest common divsor is also helpful in many theorems.
\begin{definition}[Greatest Common Divisor]
    For $f(x),g(x)\in\Qx$, we defince $\gcd(f(x),g(x)) = d(x)$ to satisfy:
    \begin{itemize}
        \item $d(x) \mid f(x)$ and $d(x) \mid g(x)$
        \item $\forall h(x) \in \Qx$ s.t. $h(x) \mid f(x)$ and $h(x) \mid g(x)$, $\|h(x)\| \leq \|d(x)\|$
    \end{itemize}
\end{definition}

\begin{theorem}[Existence and uniqueness of gcd]
Let $f(x),g(x)\in\Qx$, not both zero. Then there exists a unique monic
polynomial $d(x)\in\Qx$ such that
\[
d(x)=\gcd(f(x),g(x)).
\]
\end{theorem}

\begin{proof}
Let
\[
S=\{\,a(x)f(x)+b(x)g(x):a(x),b(x)\in\Qx,\ a(x)f(x)+b(x)g(x)\neq0\,\}.
\]
Since $f(x)$ and $g(x)$ are not both zero, the set $S$ is nonempty. Also, every
element of $S$ has positive norm. So the set
\[
N=\{\,\|s(x)\|:s(x)\in S\,\}
\]
is a nonempty subset of $\mathbb{Z}^+$. By the well-ordering principle, $N$ has
a least element. Choose $d(x)\in S$ with smallest norm. Then for some
$a(x),b(x)\in\Qx$,
\[
d(x)=a(x)f(x)+b(x)g(x).
\]

We first show that $d(x)\mid f(x)$. By the division algorithm, there exist
$q(x),r(x)\in\Qx$ such that
\[
f(x)=q(x)d(x)+r(x),
\]
where either $r(x)=0$ or
\[
\|r(x)\|<\|d(x)\|.
\]
Rearranging gives
\[
r(x)=f(x)-q(x)d(x).
\]
Since $d(x)=a(x)f(x)+b(x)g(x)$, we get
\[
r(x)=f(x)-q(x)(a(x)f(x)+b(x)g(x)).
\]
Thus
\[
r(x)=(1-q(x)a(x))f(x)+(-q(x)b(x))g(x).
\]
So if $r(x)\neq0$, then $r(x)\in S$, but it has smaller norm than $d(x)$,
which contradicts how $d(x)$ was chosen. Therefore $r(x)=0$, and hence
\[
d(x)\mid f(x).
\]
The same argument with $g(x)$ in place of $f(x)$ shows that
\[
d(x)\mid g(x).
\]

Now let $h(x)\in\Qx$ be any common divisor of $f(x)$ and $g(x)$. Then
\[
f(x)=h(x)u(x)
\qquad\text{and}\qquad
g(x)=h(x)v(x)
\]
for some $u(x),v(x)\in\Qx$. Hence
\[
d(x)=a(x)f(x)+b(x)g(x)
=a(x)h(x)u(x)+b(x)h(x)v(x).
\]
Factoring gives
\[
d(x)=h(x)(a(x)u(x)+b(x)v(x)).
\]
Thus
\[
h(x)\mid d(x).
\]
So $d(x)$ satisfies the gcd property.

Finally, make $d(x)$ monic. Let $c\in\mathbb{Q}$ be the leading coefficient of
$d(x)$. Since $d(x)\neq0$, we have $c\neq0$. Define
\[
d_0(x)=c^{-1}d(x).
\]
Then $d_0(x)$ is monic. Since $c$ is a nonzero rational number, multiplying by
$c$ or $c^{-1}$ does not change divisibility in $\Qx$. Therefore $d_0(x)$ is
also a greatest common divisor of $f(x)$ and $g(x)$.

It remains to prove uniqueness. Suppose $d_1(x)$ and $d_2(x)$ are both monic
greatest common divisors of $f(x)$ and $g(x)$. Since $d_1(x)$ is a gcd and
$d_2(x)$ is a common divisor, we have
\[
d_2(x)\mid d_1(x).
\]
Similarly,
\[
d_1(x)\mid d_2(x).
\]
So there exist $u(x),v(x)\in\Qx$ such that
\[
d_1(x)=d_2(x)u(x)
\qquad\text{and}\qquad
d_2(x)=d_1(x)v(x).
\]
Taking degrees gives
\[
\deg(d_1)=\deg(d_2)+\deg(u)
\]
and
\[
\deg(d_2)=\deg(d_1)+\deg(v).
\]
Thus $\deg(u)=\deg(v)=0$, so $u(x)$ and $v(x)$ are nonzero constant
polynomials. Since $d_1(x)$ and $d_2(x)$ are both monic, the constant $u(x)$
must be $1$. Therefore
\[
d_1(x)=d_2(x).
\]
Hence the monic greatest common divisor exists and is unique.
\end{proof}
